Quantifying the Importance of Change for Understanding Differences in Personal Culture

### 2024-08-22 write-ups/manuscript/manuscript.Rmd
adding code; changing repo structure.

In this paper, we build on the LCAM by outlining a single metric that combines components of the LCAM to quantify the relative importance of intrapersonal change and interpersonal difference for explaining variation in a population for a variable over time. Our approach isolates the "systematic" components of variance (i.e., it removes random change and measurement error), to provide a single succinct summary of how much systematic variation can reasonably be attributed to people making persistent changes over time. This quantity can then be used as a quick diagnostic to compare across questions, groups, periods, or samples to isolate where within-person change is more or less useful for explaining variation. We demonstrate the utility of this quantity in two ways.

@@

Second, to illustrate how this quantity might be used by a researcher with a specific topical interest, we investigate differences in the relative importance of intrapersonal change for people with and without a college degree. We find that change explains a smaller proportion of systematic variance among college-educated respondents, suggesting that college crystallizes one's personal culture rather than fostering an openness to new information. This offers a new empirical basis to theorize about the role of education in influencing personal culture.

>>>>

In this paper, we build on the LCAM by outlining a single metric that combines components of the LCAM to quantify the relative importance of intrapersonal change and interpersonal difference for explaining variation in a population for a variable over time. Our approach isolates the "systematic" components of variance (i.e., it removes random change and measurement error), to provide a single summary of how much systematic variation can reasonably be attributed to people making persistent changes over time. This quantity can then be used as a quick diagnostic to compare across questions, groups, periods, or samples to isolate where within-person change is more or less useful for explaining variation. We demonstrate the utility of this quantity in two ways.

@@

Second, to illustrate how this quantity might be used by a researcher with a specific topical or theoretical interest, we investigate differences in the relative importance of intrapersonal change for people with and without a college degree. We find that change explains a smaller proportion of systematic variance among college-educated respondents, suggesting that college crystallizes one's personal culture rather than fostering an openness to new information. This offers a new empirical basis to theorize about the role of education in influencing personal culture.
